---
title: "DALY Burden Estimation Workflow in anumaan"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{DALY Burden Estimation Workflow in anumaan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = TRUE,
  message  = FALSE,
  warning  = FALSE
)
```

```{r libraries}
library(anumaan)
library(dplyr)
```

# Overview

This vignette covers the **DALY burden estimation pipeline**. It does not
repeat preprocessing — that is in the
[Preprocessing Workflow vignette](preprocessing-workflow.html).

For Bayesian multivariate probit resistance-profile estimation, `anumaan`
supports both CPU and OpenCL execution backends through the `compute` argument
to `fit_bayesian_multivariate_probit()`. This is a computational choice only:
it does not alter the statistical model, validation framework, or DALY
estimands.

The pipeline has two stages:

1. **Resistance profile estimation** — converts resistance prevalence data
   into probability distributions over all 2^n binary resistance profiles
   (S/R per antibiotic class). Uses convex optimisation (GBD eq. 7.5.1.3).
2. **Burden calculation** — applies relative-risk weights to profiles and
   computes YLL, YLD, and total DALY burden.

Input can come from two sources:

| Source | Entry point |
|---|---|
| Pre-computed aggregate marginals (GBD, GLASS, national surveillance) | `validate_aggregate_inputs()` → `estimate_profiles_convex()` |
| Facility line-list AST data (after preprocessing) | `compute_marginal_resistance()` → `compute_pairwise_coresistance()` → `compute_resistance_profiles()` |

---

# Resistance Profile Estimation

## Option 1: Convex Optimisation (Pathway 1)

### 1.1 Aggregate Input: Validate Marginals

When working from pre-computed marginal resistance rates (e.g. GBD ST-GPR
country estimates or national surveillance summaries), supply them directly as
a tibble and validate before running the QP.

```{r aggregate-marginals}
marginals <- tibble::tibble(
  pathogen         = rep(c("Klebsiella pneumoniae", "Escherichia coli"), each = 3),
  antibiotic_class = rep(
    c("Carbapenems", "3GC", "Fluoroquinolones"),
    times = 2
  ),
  n_tested    = c(420L, 460L, 390L, 280L, 310L, 265L),
  n_resistant = c(126L, 299L, 195L,  45L, 155L, 133L)
) %>%
  dplyr::mutate(marginal_resistance = n_resistant / n_tested)

marginals
```

```{r validate-aggregate}
validate_aggregate_inputs(
  marginals,
  pathogen_col    = "pathogen",
  class_col       = "antibiotic_class",
  rate_col        = "marginal_resistance",
  n_tested_col    = "n_tested",
  n_resistant_col = "n_resistant"
)
```

### 1.2 Profile Enumeration

`enumerate_binary_profiles()` generates all 2^n binary combinations for a
given ordered class set. This is the profile space the QP optimises over.

```{r enumerate-profiles}
classes  <- c("Carbapenems", "3GC", "Fluoroquinolones")
profiles <- enumerate_binary_profiles(classes)
profiles
```

Each row is one resistance phenotype. `SSS` = pan-susceptible reference;
`RRR` = pan-resistant. The ordered `classes` vector defines the bit positions
and must stay consistent throughout the pipeline.

### 1.3 Constraint Matrix

`build_constraint_matrix()` constructs the constraint matrix M and target
vector v that encode the marginal and pairwise constraints fed into the QP.

```{r constraint-matrix}
kp      <- marginals[marginals$pathogen == "Klebsiella pneumoniae", ]
r_marg  <- setNames(kp$marginal_resistance, kp$antibiotic_class)
profiles_enum <- enumerate_binary_profiles(names(r_marg))
cm      <- build_constraint_matrix(profiles_enum, r_marg)

cat("M dimension (constraints x profiles):", dim(cm$M), "\n")
cat("Constraint targets (marginals + pairwise independence fallback):\n")
round(cm$v, 4)
```

Marginal rows of M: entry = 1 if that class is resistant in that profile.
Pairwise rows: entry = 1 if both classes in the pair are resistant.

### 1.4 Estimate Profile Probabilities

`estimate_profiles_convex()` runs the full pipeline for all pathogens in one
call: validates, enumerates, builds constraints, and solves the QP. The solver
prefers `osqp` (sparse, fast) with `quadprog` as fallback.

```{r estimate-profiles}
panel_map <- list(
  "Klebsiella pneumoniae"  = c("Carbapenems",
                                "3GC",
                                "Fluoroquinolones"),
  "Escherichia coli"       = c("Carbapenems",
                                "3GC",
                                "Fluoroquinolones")
)

profiles_out <- estimate_profiles_convex(
  marginals    = marginals,
  pairwise     = NULL,
  panel_map    = panel_map,
  lambda       = 1e-8,
  pathogen_col = "pathogen",
  class_col    = "antibiotic_class",
  rate_col     = "marginal_resistance",
  n_tested_col = "n_tested"
)
```

```{r show-profiles}
profiles_out %>%
  dplyr::select(pathogen, profile_delta, profile_probability,
                convergence_flag, max_abs_residual) %>%
  dplyr::filter(profile_probability > 0.005) %>%
  dplyr::arrange(pathogen, dplyr::desc(profile_probability))
```

The `profile_class_set` column records the exact ordered class set that
defines the binary code — this must be carried forward to any downstream
DALY attribution to prevent class-order ambiguity.

```{r profile-class-set}
profiles_out %>%
  dplyr::distinct(pathogen, profile_class_set, estimator)
```

### 1.5 Facility Line-List Pathway

When working from isolate-level AST data (after preprocessing), use the
three-step pipeline that feeds the same QP engine.

```{r line-list-data}
set.seed(101)
# Minimal long-format AST data: one row per isolate x antibiotic class
ast_class <- data.frame(
  isolate_id       = rep(paste0("ISO", sprintf("%03d", 1:80)), each = 3),
  organism_name    = rep(
    ifelse(seq_len(80) <= 50, "Klebsiella pneumoniae", "Escherichia coli"),
    each = 3
  ),
  org_group        = "Enterobacterales",
  antibiotic_class = rep(c("Carbapenems", "3GC", "Fluoroquinolones"), times = 80),
  antibiotic_value = sample(
    c("S", "R"), 240, replace = TRUE, prob = c(0.60, 0.40)
  ),
  stringsAsFactors = FALSE
)
```

```{r step1-marginal}
# Step 1 -- marginal resistance per pathogen x class
marg_out <- compute_marginal_resistance(
  ast_class,
  pathogen_col         = "organism_name",
  org_group_col        = "org_group",
  isolate_col          = "isolate_id",
  antibiotic_class_col = "antibiotic_class",
  antibiotic_value_col = "antibiotic_value",
  min_n_tested         = 10L
)

marg_out$marginal
```

```{r step2-pairwise}
# Step 2 -- pairwise co-resistance matrices per pathogen
co_out <- compute_pairwise_coresistance(
  marg_out,
  pathogen_col         = "organism_name",
  isolate_col          = "isolate_id",
  antibiotic_class_col = "antibiotic_class",
  min_co_tested        = 5L
)

round(co_out[["Klebsiella pneumoniae"]]$prevalence, 3)
```

```{r step3-profiles}
# Step 3 -- profile probabilities via QP
rp_out <- compute_resistance_profiles(
  marg_out,
  co_out,
  pathogen_col         = "organism_name",
  antibiotic_class_col = "antibiotic_class",
  exclude_near_zero    = FALSE
)

rp_out[["Klebsiella pneumoniae"]]$profiles %>%
  dplyr::filter(probability > 0.005) %>%
  dplyr::arrange(dplyr::desc(probability))
```

```{r constraint-residuals}
# Constraint residuals: how well does the solution reproduce the inputs?
round(rp_out[["Klebsiella pneumoniae"]]$constraint_residuals, 6)
```

### 1.6 Check Constraint Satisfaction

`check_profile_constraints()` formally verifies that the estimated
probabilities satisfy non-negativity, sum-to-one, and reproduce the input
marginal rates within tolerance. It accepts the named-list format from
`compute_resistance_profiles()` directly.

```{r check-constraints}
checks <- check_profile_constraints(
  rp_out,
  marginals    = marg_out$marginal,
  tolerance    = 1e-3,
  pathogen_col = "organism_name",
  class_col    = "antibiotic_class",
  rate_col     = "marginal_resistance"
)

checks %>%
  dplyr::select(pathogen, constraint_type, constraint_name,
                target, reconstructed, abs_residual, pass)
```

### 1.7 Bootstrap Uncertainty Intervals

`bootstrap_profiles_convex()` resamples resistant counts from a Binomial
distribution and refits the QP B times, returning percentile confidence
intervals for each profile probability.

```{r bootstrap-profiles}
boot <- bootstrap_profiles_convex(
  marginals       = marginals,
  B               = 300L,
  seed            = 42L,
  alpha           = 0.05,
  pathogen_col    = "pathogen",
  class_col       = "antibiotic_class",
  n_tested_col    = "n_tested",
  n_resistant_col = "n_resistant"
)

boot[["Klebsiella pneumoniae"]] %>%
  dplyr::filter(probability_mean > 0.005) %>%
  dplyr::arrange(dplyr::desc(probability_mean))
```

## Option 2: Bayesian Multivariate Probit (Pathway 2)

Pathway 2 fits a Bayesian hierarchical multivariate probit model directly to
facility-level, event-level AST data. Unlike Pathway 1, it can incorporate
fixed-effect covariates (age, gender, location, ...) and optional random
effects (hospital, admission, ...), and estimate resistance *jointly* across
antibiotic classes rather than reconstructing joint profiles purely from
marginal + pairwise constraints.

This requires the optional `cmdstanr` + CmdStan dependency (see
`DESCRIPTION`) and fits an actual MCMC model, so the code chunks below are
shown with `eval = FALSE` for portability -- the printed output is real,
captured from an actual run on the synthetic data below (2 chains, 200
warmup/200 sampling iterations -- deliberately small and fast for
illustration; a real analysis uses far more, e.g. 4 chains x 3000
warmup/1000 sampling).

### 2.1 Prepare Event-Level Class Data

One row per organism-event. Antibiotic class columns hold `0` (susceptible),
`1` (resistant), or `NA` (not tested) -- a different encoding from Pathway
1's S/R text values, because this is the direct input to the Stan model.

```{r probit-data, eval = FALSE}
set.seed(123)
n_events <- 150
centers  <- c("Hospital A", "Hospital B", "Hospital C")

event_class_data <- tibble::tibble(
  event_id       = paste0("EV", sprintf("%04d", 1:n_events)),
  pathogen       = "Klebsiella pneumoniae",
  center_name    = sample(centers, n_events, replace = TRUE),
  Age_normalised = round(runif(n_events, 0, 90), 1),
  gender         = sample(c("Male", "Female"), n_events, replace = TRUE),
  final_outcome  = sample(c("Discharged", "Died"), n_events, replace = TRUE, prob = c(0.8, 0.2))
)

true_p <- c(Carbapenems = 0.35, Fluoroquinolones = 0.55, Aminoglycosides = 0.30)
for (cls in names(true_p)) {
  vals <- rbinom(n_events, 1, true_p[[cls]])
  vals[sample.int(n_events, size = floor(0.15 * n_events))] <- NA  # some untested
  event_class_data[[cls]] <- vals
}

class_cols <- c("Carbapenems", "Fluoroquinolones", "Aminoglycosides")
```

### 2.2 Fit the Model

`fixed_effects` is required. `random_effects` may be a legacy character vector,
a named list-of-blocks specification, or `list()` for a fixed-effects-only
model. When no random effects are fitted, `profile_group_col` is required to
state the column used for profile aggregation and validation; it does not make
that column a random effect. `residual_structure = "identity"` treats classes
as conditionally independent given the modelled mean (default, more stable);
`"correlated"` estimates a full residual correlation matrix via an
LKJCholesky prior, but needs adequate pairwise co-testing overlap to be
identifiable (`fit$eligibility_report$pairwise`).

```{r probit-fit, eval = FALSE}
fit <- fit_bayesian_multivariate_probit(
  event_class_data   = event_class_data,
  class_cols         = class_cols,
  fixed_effects      = c("Age_normalised", "gender"),
  random_effects     = c("center_name"),
  pathogen           = "Klebsiella pneumoniae",
  outcome_col        = "final_outcome",
  residual_structure = "identity",
  prior_config       = list(beta_sd = 1.5, tau_sd = 1.0),
  sampler_config     = list(chains = 2, iter_warmup = 200, iter_sampling = 200,
                             seed = 123, parallel_chains = 2, adapt_delta = 0.9)
)

fit$diagnostics
#> # A tibble: 1 x 36
#>   n_chains iter_warmup iter_sampling n_re_levels n_observed_pairs n_events
#>      <int>       <int>         <int>       <int>            <int>    <int>
#> 1        2         200           200           1              384      149
#> # i 30 more variables: n_classes <int>, max_rhat_structural <dbl>,
#> #   min_ess_bulk_structural <dbl>, min_ess_tail_structural <dbl>,
#> #   n_divergent <int>, converged_structural <lgl>, diagnostic_status <chr>, ...
```

For a fixed-effects-only comparison, retain the same fixed-effect design and
make the grouping purpose explicit without adding a random-effect block:

```r
fixed_only_fit <- fit_bayesian_multivariate_probit(
  event_class_data = event_class_data,
  class_cols = class_cols,
  fixed_effects = c("Age_normalised", "gender", "center_name"),
  random_effects = list(),
  profile_group_col = "center_name",
  pathogen = "Klebsiella pneumoniae",
  residual_structure = "identity"
)
```

With these deliberately tiny/fast settings, `converged_structural` comes
back `FALSE` (max R-hat ~1.03, a few divergences) -- expected for a 200/200
smoke-test fit, not a sign of a broken model. This mirrors the
`exp_00_smoke_test` convention used for fast iteration before committing to
a full run with production-scale `sampler_config`.

### 2.3 Convert Posterior Draws to Resistance-Profile Probabilities

```{r probit-profiles, eval = FALSE}
profiles <- compute_event_profile_probabilities(
  fit,
  n_posterior_draws_for_profiles = 200L,
  outcome_col = "final_outcome",
  seed = 123L
)

names(profiles)
#> [1] "event_profiles"  "aggregate_draws"

profiles$event_profiles
#> [compute_event_profile_probabilities] 200 draws | 149 events | 3 hp-pairs | 1 RE level(s)
#> # A tibble: 1,192 x 14
#>   center_name pathogen               event_idx profile_class_set   profile_delta
#>   <chr>       <chr>                      <int> <chr>               <chr>
#> 1 Hospital C  Klebsiella pneumoniae         1   Carbapenems|Fluoro… SSS
#> 2 Hospital C  Klebsiella pneumoniae         1   Carbapenems|Fluoro… RSS
#> # i 1,190 more rows, and 9 more variables: profile_probability <dbl>, ...
```

`event_profiles` is one row per event x profile (up to 2^D rows per event);
`aggregate_draws` carries per-draw `R_ALL`/`R_KNOWN_OUTCOME`/`R_NF` used for
credible intervals in the next step.

### 2.4 Aggregate for DALY

```{r probit-aggregate, eval = FALSE}
agg <- aggregate_profiles_for_daly(profiles, hospital_col = "center_name", pathogen_col = "pathogen")

agg
#> [aggregate_profiles_for_daly] 24 hospital-pathogen-profile rows
#> # A tibble: 24 x 44
#>   center_name pathogen        profile_class_set    profile_delta R_ALL_mean R_ALL_lower R_ALL_upper ...
#>   <chr>       <chr>           <chr>                <chr>              <dbl>       <dbl>       <dbl>
#> 1 Hospital A  Klebsiella pne… Carbapenems|Fluoro…   RRR
#> # i 23 more rows
```

`agg` is analogous to Pathway 1's `profiles_out`/`rp_out` -- both are the
resistance-profile output this vignette produces; downstream burden
calculation (`daly_assign_rr_to_profiles()` onward) is not yet covered here.

### 2.5 Validate Calibration

Three calibration checks compare the model's posterior predictions back
against the observed data it was fit on, at increasing levels of joint
complexity: single-class marginals, pairs of classes, and complete observed
profiles.

```{r probit-validate, eval = FALSE}
val_marginal <- validate_marginal_calibration(fit, n_posterior_draws_for_validation = 200L, seed = 123L)
val_pairwise <- validate_pairwise_calibration(fit, n_posterior_draws_for_validation = 200L, seed = 123L)
val_complete <- validate_complete_profile_calibration(fit, n_posterior_draws_for_validation = 200L,
                                                       seed = 123L, min_complete_events = 5L)

compute_profile_validation_status(
  marginal_tbl          = val_marginal,
  pairwise_tbl          = val_pairwise,
  complete_profile_tbl  = val_complete
)
#> $status
#> [1] "pass"
#> $reasons
#> character(0)
#> $thresholds_used
#> $thresholds_used$max_mean_abs_error_marginal
#> [1] 0.1
#> ...
```

`compute_profile_validation_status()` is the single authoritative pass/fail
call -- it applies default thresholds (mean absolute error, interval
coverage) across all three tables and returns one combined `status`, rather
than requiring the caller to eyeball three separate tibbles.

### 2.6 Predictive Checks

Beyond calibration against the fitted data, prior and posterior predictive
checks simulate data *from* the model to check the priors are reasonable
before fitting, and that the fitted model reproduces the observed data's
statistical structure after fitting.

```{r probit-predictive, eval = FALSE}
prior_pred <- simulate_probit_prior_predictive(fit, n_states = 100L, seed = 123L)
compute_prior_predictive_status(prior_pred)
#> $summary$fraction_probability_lt_0.001
#> [1] 0.478
#> $summary$fraction_all_resistant
#> [1] 0.0857

ppc_stats <- compute_probit_ppc_statistics(
  fit, n_states = 100L, seed = 123L,
  statistics = c("marginal", "resistant_count", "pairwise")
)
compute_posterior_predictive_status(ppc_stats)
#> $status
#> [1] "pass"
#> $family_status$marginal$status
#> [1] "ok"
#> $family_status$pairwise$status
#> [1] "ok"
```

`plot_probit_diagnostics(fit, output_dir, experiment_id, pathogen)` (needs
`bayesplot`) writes trace/rank/pair plots for the monitored parameters to a
PDF, and `plot_probit_posterior_predictive_checks()` visualises the
`ppc_stats` tables -- both are for interactive review, not part of the
scripted pipeline.

### 2.7 Unified Dispatcher

`estimate_resistance_profiles()` wraps both pathways behind one interface --
switch pathway with `method`:

```{r probit-dispatcher, eval = FALSE}
# Pathway 1 (convex)
estimate_resistance_profiles(data = marginals, method = "convex", panel_map = panel_map)

# Pathway 2 (Bayesian) -- runs steps 2.1-2.4 above in one call
estimate_resistance_profiles(
  data = event_class_data, method = "bayesian",
  class_cols = class_cols, fixed_effects = c("Age_normalised", "gender"),
  random_effects = c("center_name"), pathogen = "Klebsiella pneumoniae",
  outcome_col = "final_outcome", residual_structure = "identity",
  sampler_config = list(chains = 2, iter_warmup = 200, iter_sampling = 200, seed = 123)
)
#> [estimate_resistance_profiles] Pathway 2 complete.
#> $profiles      # same shape as aggregate_profiles_for_daly() output above
#> $eligibility
#> $diagnostics
#> $fitted_models
#> $config_used
```

Both pathways' `profiles`/`agg` output are the deliverable of this vignette
-- downstream burden calculation (RR assignment, YLL/YLD) is not yet covered
here.

### 2.8 Post-Fit Checklist: What to Run, and Why

Sections 2.2-2.6 introduced each Pathway 2 function individually; this
section ties them into one ordered checklist, since it's easy to run only
some of them and miss a gap. Every function below reads `fit`, the object
returned by `fit_bayesian_multivariate_probit()` -- none of them require
re-fitting.

**1. Did the sampler converge?** (`fit$diagnostics`, populated at fit time --
no separate call needed). Check `diagnostic_status` (`"pass"` /
`"warning_rhat"` / `"fail_energy"` / `"fail_divergent"` / ...),
`converged_structural`, `max_rhat`, `min_ess_bulk`, `ebfmi_min`. This is a
*computational* question -- did the four chains actually agree with each
other about where the posterior mass is -- and it is answered independently
of everything below. A model can fail here while still answering later
checks well (the reverse is not reassuring: don't trust calibration numbers
from an unconverged fit).

**2. Estimate the resistance profiles.**
`compute_event_profile_probabilities()` computes, per event, the posterior
probability of every possible resistance profile -- analytically
(`Phi(mu_d)` per class) for `residual_structure = "identity"`, or via Gibbs
sampling of the truncated multivariate normal conditional on that event's
own observed AST results for `"correlated"`. It never overwrites an
observed cell -- only untested classes are imputed, so a fully-observed
event's profile is a fact about the data, not a claim from the model.
`aggregate_profiles_for_daly()` rolls the per-event, per-draw output up to
one row per hospital x pathogen x profile (`R_ALL_mean/lower/upper`,
DALY-eligibility flags gated by `sampler_acceptable` and panel support).

**3. Does the model generate data that looks like the real data?**
`simulate_probit_posterior_predictive()` draws replicated datasets from the
posterior (Bernoulli(`Phi(mu)`) for identity; `Z = mu + L_Omega %*% eps`,
`Y = I(Z > 0)` for correlated -- genuine correlated simulation either way).
`compute_probit_ppc_statistics()` computes discrepancy statistics (marginal
rate, pairwise co-resistance, complete-profile summary stats, hospital
heterogeneity) on observed vs. every replicate.
`compute_posterior_predictive_status()` classifies each statistic as
extreme/severe and rolls up to an overall status. This is a strict trigger,
not a fraction-based one -- even a single severely-extreme statistic in a
family sets the whole run's status to `"fail_major_ppc_misfit"`, so always
read the per-family `n_severe`/`fraction_extreme` breakdown underneath the
top-line status rather than the label alone. `simulate_probit_prior_predictive()`
(before fitting) and `simulate_probit_mixed_predictive()` (new, unseen
groups) answer the adjacent but different questions "is the prior
reasonable?" and "does this generalise beyond the fitted groups?".

**4. Are the model's stated probabilities calibrated against reality?**
`validate_marginal_calibration()`, `validate_pairwise_calibration()`, and
`validate_complete_profile_calibration()` (Section 2.5) compare observed
rates to model-implied probabilities at increasing joint complexity --
single class, class pairs, and full profiles. All three correctly use
`Omega` for correlated fits where relevant (pairwise and complete-profile;
marginal probabilities don't depend on Omega by construction).
`mask_and_validate_ast()` adds a masked-AST holdout check -- by default
(`refit = FALSE`) this is an *in-sample* diagnostic (the masked cells were
still seen during fitting, labelled `"in_sample_no_refit"` in its output,
not a genuine holdout), while `refit = TRUE` actually refits without the
masked cells for a real out-of-sample score, at proportionally higher cost.
`compute_profile_validation_status()` rolls all four into one
`profile_validation_status` verdict.

**Quick reference: which check answers which question**

| Question | Function | Needs `Omega`? |
|---|---|---|
| Did the sampler converge? | `fit$diagnostics` | -- |
| What's the estimated resistance profile? | `compute_event_profile_probabilities()` + `aggregate_profiles_for_daly()` | correlated fits only |
| Does simulated data resemble real data? | `simulate_probit_posterior_predictive()` + `compute_probit_ppc_statistics()` | correlated fits only |
| Is the model calibrated on single classes? | `validate_marginal_calibration()` | no |
| Is the model calibrated on class pairs? | `validate_pairwise_calibration()` | correlated fits only |
| Is the model calibrated on full profiles? | `validate_complete_profile_calibration()` | correlated fits only |
| Can it predict held-out AST results? | `mask_and_validate_ast()` | correlated fits only |

A model can score well on some of these and poorly on others at the same
time -- e.g. a correlated-residual fit with unresolved sampler convergence
issues can still show much better calibration on class pairs and full
profiles than a well-converged identity-residual fit, because convergence
and calibration are answering different questions (numerical reliability
of the specific parameter estimates vs. structural adequacy of the model
form). Treat "did it converge" and "is it calibrated" as two separate
verdicts, not one combined score.

---

# Pipeline at a Glance

```
Preprocessing output (prep_* pipeline)
          │
          ├─────────────────────────────┬─────────────────────────────────────┐
          ▼ Option 1: Convex (Pathway 1) ▼                Option 2: Bayesian Probit (Pathway 2)
  compute_marginal_resistance()          fit_bayesian_multivariate_probit()
          │  (Step 1: marginals)                 │  (facility event-level AST + covariates)
          ▼                                       ▼
  compute_pairwise_coresistance()        compute_event_profile_probabilities()
          │  (Step 2: pairwise co-R)             │  (posterior draws -> profile probs)
          ▼                                       ▼
  compute_resistance_profiles()          aggregate_profiles_for_daly()
          │  (Step 3: QP -> profiles)            │  (+ validate_*_calibration(),
          │                                       │    predictive checks)
          │    Alternative entry (aggregate       │
          │    marginals): validate_aggregate_    │
          │    inputs() -> estimate_profiles_     │
          │    convex()                           │
          │                                       │
          │         (or, either pathway via one call:
          │          estimate_resistance_profiles(method = "convex" | "bayesian"))
          │                                       │
          └───────────────────┬───────────────────┘
                               ▼
                  Resistance profile probabilities (this vignette's scope ends here;
                  downstream burden calculation -- RR assignment, YLL/YLD -- not yet covered)
```

---

# Session Info

```{r session}
sessionInfo()
```
