---
title: "DALY Burden Estimation Workflow in anumaan"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{DALY Burden Estimation Workflow in anumaan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = TRUE,
  message  = FALSE,
  warning  = FALSE
)
```

```{r libraries}
library(anumaan)
library(dplyr)
```

# Overview

This vignette covers the **DALY burden estimation pipeline**. It does not
repeat preprocessing — that is in the
[Preprocessing Workflow vignette](preprocessing-workflow.html).

For Bayesian multivariate probit resistance-profile estimation, `anumaan`
supports both CPU and OpenCL execution backends through the `compute` argument
to `fit_bayesian_multivariate_probit()`. This is a computational choice only:
it does not alter the statistical model, validation framework, or DALY
estimands.

The pipeline has two stages:

1. **Resistance profile estimation** — converts resistance prevalence data
   into probability distributions over all 2^n binary resistance profiles
   (S/R per antibiotic class). Uses convex optimisation (GBD eq. 7.5.1.3).
2. **Burden calculation** — applies relative-risk weights to profiles and
   computes YLL, YLD, and total DALY burden.

Input can come from two sources:

| Source | Entry point |
|---|---|
| Pre-computed aggregate marginals (GBD, GLASS, national surveillance) | `validate_aggregate_inputs()` → `estimate_profiles_convex()` |
| Facility line-list AST data (after preprocessing) | `compute_marginal_resistance()` → `compute_pairwise_coresistance()` → `compute_resistance_profiles()` |

---

# Resistance Profile Estimation

## Option 1: Convex Optimisation (Pathway 1)

### 1.1 Aggregate Input: Validate Marginals

When working from pre-computed marginal resistance rates (e.g. GBD ST-GPR
country estimates or national surveillance summaries), supply them directly as
a tibble and validate before running the QP.

```{r aggregate-marginals}
marginals <- tibble::tibble(
  pathogen         = rep(c("Klebsiella pneumoniae", "Escherichia coli"), each = 3),
  antibiotic_class = rep(
    c("Carbapenems", "3GC", "Fluoroquinolones"),
    times = 2
  ),
  n_tested    = c(420L, 460L, 390L, 280L, 310L, 265L),
  n_resistant = c(126L, 299L, 195L,  45L, 155L, 133L)
) %>%
  dplyr::mutate(marginal_resistance = n_resistant / n_tested)

marginals
```

```{r validate-aggregate}
validate_aggregate_inputs(
  marginals,
  pathogen_col    = "pathogen",
  class_col       = "antibiotic_class",
  rate_col        = "marginal_resistance",
  n_tested_col    = "n_tested",
  n_resistant_col = "n_resistant"
)
```

### 1.2 Profile Enumeration

`enumerate_binary_profiles()` generates all 2^n binary combinations for a
given ordered class set. This is the profile space the QP optimises over.

```{r enumerate-profiles}
classes  <- c("Carbapenems", "3GC", "Fluoroquinolones")
profiles <- enumerate_binary_profiles(classes)
profiles
```

Each row is one resistance phenotype. `SSS` = pan-susceptible reference;
`RRR` = pan-resistant. The ordered `classes` vector defines the bit positions
and must stay consistent throughout the pipeline.

### 1.3 Constraint Matrix

`build_constraint_matrix()` constructs the constraint matrix M and target
vector v that encode the marginal and pairwise constraints fed into the QP.

```{r constraint-matrix}
kp      <- marginals[marginals$pathogen == "Klebsiella pneumoniae", ]
r_marg  <- setNames(kp$marginal_resistance, kp$antibiotic_class)
profiles_enum <- enumerate_binary_profiles(names(r_marg))
cm      <- build_constraint_matrix(profiles_enum, r_marg)

cat("M dimension (constraints x profiles):", dim(cm$M), "\n")
cat("Constraint targets (marginals + pairwise independence fallback):\n")
round(cm$v, 4)
```

Marginal rows of M: entry = 1 if that class is resistant in that profile.
Pairwise rows: entry = 1 if both classes in the pair are resistant.

### 1.4 Estimate Profile Probabilities

`estimate_profiles_convex()` runs the full pipeline for all pathogens in one
call: validates, enumerates, builds constraints, and solves the QP. The solver
prefers `osqp` (sparse, fast) with `quadprog` as fallback.

```{r estimate-profiles}
panel_map <- list(
  "Klebsiella pneumoniae"  = c("Carbapenems",
                                "3GC",
                                "Fluoroquinolones"),
  "Escherichia coli"       = c("Carbapenems",
                                "3GC",
                                "Fluoroquinolones")
)

profiles_out <- estimate_profiles_convex(
  marginals    = marginals,
  pairwise     = NULL,
  panel_map    = panel_map,
  lambda       = 1e-8,
  pathogen_col = "pathogen",
  class_col    = "antibiotic_class",
  rate_col     = "marginal_resistance",
  n_tested_col = "n_tested"
)
```

```{r show-profiles}
profiles_out %>%
  dplyr::select(pathogen, profile_delta, profile_probability,
                convergence_flag, max_abs_residual) %>%
  dplyr::filter(profile_probability > 0.005) %>%
  dplyr::arrange(pathogen, dplyr::desc(profile_probability))
```

The `profile_class_set` column records the exact ordered class set that
defines the binary code — this must be carried forward to any downstream
DALY attribution to prevent class-order ambiguity.

```{r profile-class-set}
profiles_out %>%
  dplyr::distinct(pathogen, profile_class_set, estimator)
```

### 1.5 Facility Line-List Pathway

When working from isolate-level AST data (after preprocessing), use the
three-step pipeline that feeds the same QP engine.

```{r line-list-data}
set.seed(101)
# Minimal long-format AST data: one row per isolate x antibiotic class
ast_class <- data.frame(
  isolate_id       = rep(paste0("ISO", sprintf("%03d", 1:80)), each = 3),
  organism_name    = rep(
    ifelse(seq_len(80) <= 50, "Klebsiella pneumoniae", "Escherichia coli"),
    each = 3
  ),
  org_group        = "Enterobacterales",
  antibiotic_class = rep(c("Carbapenems", "3GC", "Fluoroquinolones"), times = 80),
  antibiotic_value = sample(
    c("S", "R"), 240, replace = TRUE, prob = c(0.60, 0.40)
  ),
  stringsAsFactors = FALSE
)
```

```{r step1-marginal}
# Step 1 -- marginal resistance per pathogen x class
marg_out <- compute_marginal_resistance(
  ast_class,
  pathogen_col         = "organism_name",
  org_group_col        = "org_group",
  isolate_col          = "isolate_id",
  antibiotic_class_col = "antibiotic_class",
  antibiotic_value_col = "antibiotic_value",
  min_n_tested         = 10L
)

marg_out$marginal
```

```{r step2-pairwise}
# Step 2 -- pairwise co-resistance matrices per pathogen
co_out <- compute_pairwise_coresistance(
  marg_out,
  pathogen_col         = "organism_name",
  isolate_col          = "isolate_id",
  antibiotic_class_col = "antibiotic_class",
  min_co_tested        = 5L
)

round(co_out[["Klebsiella pneumoniae"]]$prevalence, 3)
```

```{r step3-profiles}
# Step 3 -- profile probabilities via QP
rp_out <- compute_resistance_profiles(
  marg_out,
  co_out,
  pathogen_col         = "organism_name",
  antibiotic_class_col = "antibiotic_class",
  exclude_near_zero    = FALSE
)

rp_out[["Klebsiella pneumoniae"]]$profiles %>%
  dplyr::filter(probability > 0.005) %>%
  dplyr::arrange(dplyr::desc(probability))
```

```{r constraint-residuals}
# Constraint residuals: how well does the solution reproduce the inputs?
round(rp_out[["Klebsiella pneumoniae"]]$constraint_residuals, 6)
```

### 1.6 Check Constraint Satisfaction

`check_profile_constraints()` formally verifies that the estimated
probabilities satisfy non-negativity, sum-to-one, and reproduce the input
marginal rates within tolerance. It accepts the named-list format from
`compute_resistance_profiles()` directly.

```{r check-constraints}
checks <- check_profile_constraints(
  rp_out,
  marginals    = marg_out$marginal,
  tolerance    = 1e-3,
  pathogen_col = "organism_name",
  class_col    = "antibiotic_class",
  rate_col     = "marginal_resistance"
)

checks %>%
  dplyr::select(pathogen, constraint_type, constraint_name,
                target, reconstructed, abs_residual, pass)
```

### 1.7 Bootstrap Uncertainty Intervals

`bootstrap_profiles_convex()` resamples resistant counts from a Binomial
distribution and refits the QP B times, returning percentile confidence
intervals for each profile probability.

```{r bootstrap-profiles}
boot <- bootstrap_profiles_convex(
  marginals       = marginals,
  B               = 300L,
  seed            = 42L,
  alpha           = 0.05,
  pathogen_col    = "pathogen",
  class_col       = "antibiotic_class",
  n_tested_col    = "n_tested",
  n_resistant_col = "n_resistant"
)

boot[["Klebsiella pneumoniae"]] %>%
  dplyr::filter(probability_mean > 0.005) %>%
  dplyr::arrange(dplyr::desc(probability_mean))
```

## Option 2: Bayesian Multivariate Probit (Pathway 2)

Pathway 2 fits a Bayesian hierarchical multivariate probit model directly to
facility-level, event-level AST data. Unlike Pathway 1, it can incorporate
fixed-effect covariates (age, gender, location, ...) and random effects
(hospital, admission, ...), and estimate resistance *jointly* across
antibiotic classes rather than reconstructing joint profiles purely from
marginal + pairwise constraints.

This requires the optional `cmdstanr` + CmdStan dependency (see
`DESCRIPTION`) and fits an actual MCMC model, so the code chunks below are
shown with `eval = FALSE` for portability -- the printed output is real,
captured from an actual run on the synthetic data below (2 chains, 200
warmup/200 sampling iterations -- deliberately small and fast for
illustration; a real analysis uses far more, e.g. 4 chains x 3000
warmup/1000 sampling).

### 2.1 Prepare Event-Level Class Data

One row per organism-event. Antibiotic class columns hold `0` (susceptible),
`1` (resistant), or `NA` (not tested) -- a different encoding from Pathway
1's S/R text values, because this is the direct input to the Stan model.

```{r probit-data, eval = FALSE}
set.seed(123)
n_events <- 150
centers  <- c("Hospital A", "Hospital B", "Hospital C")

event_class_data <- tibble::tibble(
  event_id       = paste0("EV", sprintf("%04d", 1:n_events)),
  pathogen       = "Klebsiella pneumoniae",
  center_name    = sample(centers, n_events, replace = TRUE),
  Age_normalised = round(runif(n_events, 0, 90), 1),
  gender         = sample(c("Male", "Female"), n_events, replace = TRUE),
  final_outcome  = sample(c("Discharged", "Died"), n_events, replace = TRUE, prob = c(0.8, 0.2))
)

true_p <- c(Carbapenems = 0.35, Fluoroquinolones = 0.55, Aminoglycosides = 0.30)
for (cls in names(true_p)) {
  vals <- rbinom(n_events, 1, true_p[[cls]])
  vals[sample.int(n_events, size = floor(0.15 * n_events))] <- NA  # some untested
  event_class_data[[cls]] <- vals
}

class_cols <- c("Carbapenems", "Fluoroquinolones", "Aminoglycosides")
```

### 2.2 Fit the Model

`fixed_effects` and `random_effects` are required, with no defaults --
`random_effects` accepts 1-3 grouping columns (hospital; [+patient];
[+admission]), or the generic list-of-blocks spec for more than three.
`residual_structure = "identity"` treats classes as conditionally
independent given the fixed/random effects (default, more stable);
`"correlated"` estimates a full residual correlation matrix via an
LKJCholesky prior, but needs adequate pairwise co-testing overlap to be
identifiable (`fit$eligibility_report$pairwise`).

```{r probit-fit, eval = FALSE}
fit <- fit_bayesian_multivariate_probit(
  event_class_data   = event_class_data,
  class_cols         = class_cols,
  fixed_effects      = c("Age_normalised", "gender"),
  random_effects     = c("center_name"),
  pathogen           = "Klebsiella pneumoniae",
  outcome_col        = "final_outcome",
  residual_structure = "identity",
  prior_config       = list(beta_sd = 1.5, tau_sd = 1.0),
  sampler_config     = list(chains = 2, iter_warmup = 200, iter_sampling = 200,
                             seed = 123, parallel_chains = 2, adapt_delta = 0.9)
)

fit$diagnostics
#> # A tibble: 1 x 36
#>   n_chains iter_warmup iter_sampling n_re_levels n_observed_pairs n_events
#>      <int>       <int>         <int>       <int>            <int>    <int>
#> 1        2         200           200           1              384      149
#> # i 30 more variables: n_classes <int>, max_rhat_structural <dbl>,
#> #   min_ess_bulk_structural <dbl>, min_ess_tail_structural <dbl>,
#> #   n_divergent <int>, converged_structural <lgl>, diagnostic_status <chr>, ...
```

With these deliberately tiny/fast settings, `converged_structural` comes
back `FALSE` (max R-hat ~1.03, a few divergences) -- expected for a 200/200
smoke-test fit, not a sign of a broken model. This mirrors the
`exp_00_smoke_test` convention used for fast iteration before committing to
a full run with production-scale `sampler_config`.

### 2.3 Convert Posterior Draws to Resistance-Profile Probabilities

```{r probit-profiles, eval = FALSE}
profiles <- compute_event_profile_probabilities(
  fit,
  n_posterior_draws_for_profiles = 200L,
  outcome_col = "final_outcome",
  seed = 123L
)

names(profiles)
#> [1] "event_profiles"  "aggregate_draws"

profiles$event_profiles
#> [compute_event_profile_probabilities] 200 draws | 149 events | 3 hp-pairs | 1 RE level(s)
#> # A tibble: 1,192 x 14
#>   center_name pathogen               event_idx profile_class_set   profile_delta
#>   <chr>       <chr>                      <int> <chr>               <chr>
#> 1 Hospital C  Klebsiella pneumoniae         1   Carbapenems|Fluoro… SSS
#> 2 Hospital C  Klebsiella pneumoniae         1   Carbapenems|Fluoro… RSS
#> # i 1,190 more rows, and 9 more variables: profile_probability <dbl>, ...
```

`event_profiles` is one row per event x profile (up to 2^D rows per event);
`aggregate_draws` carries per-draw `R_ALL`/`R_KNOWN_OUTCOME`/`R_NF` used for
credible intervals in the next step.

### 2.4 Aggregate for DALY

```{r probit-aggregate, eval = FALSE}
agg <- aggregate_profiles_for_daly(profiles, hospital_col = "center_name", pathogen_col = "pathogen")

agg
#> [aggregate_profiles_for_daly] 24 hospital-pathogen-profile rows
#> # A tibble: 24 x 44
#>   center_name pathogen        profile_class_set    profile_delta R_ALL_mean R_ALL_lower R_ALL_upper ...
#>   <chr>       <chr>           <chr>                <chr>              <dbl>       <dbl>       <dbl>
#> 1 Hospital A  Klebsiella pne… Carbapenems|Fluoro…   RRR
#> # i 23 more rows
```

`agg` is analogous to Pathway 1's `profiles_out`/`rp_out` -- both are the
resistance-profile output this vignette produces; downstream burden
calculation (`daly_assign_rr_to_profiles()` onward) is not yet covered here.

### 2.5 Validate Calibration

Three calibration checks compare the model's posterior predictions back
against the observed data it was fit on, at increasing levels of joint
complexity: single-class marginals, pairs of classes, and complete observed
profiles.

```{r probit-validate, eval = FALSE}
val_marginal <- validate_marginal_calibration(fit, n_posterior_draws_for_validation = 200L, seed = 123L)
val_pairwise <- validate_pairwise_calibration(fit, n_posterior_draws_for_validation = 200L, seed = 123L)
val_complete <- validate_complete_profile_calibration(fit, n_posterior_draws_for_validation = 200L,
                                                       seed = 123L, min_complete_events = 5L)

compute_profile_validation_status(
  marginal_tbl          = val_marginal,
  pairwise_tbl          = val_pairwise,
  complete_profile_tbl  = val_complete
)
#> $status
#> [1] "pass"
#> $reasons
#> character(0)
#> $thresholds_used
#> $thresholds_used$max_mean_abs_error_marginal
#> [1] 0.1
#> ...
```

`compute_profile_validation_status()` is the single authoritative pass/fail
call -- it applies default thresholds (mean absolute error, interval
coverage) across all three tables and returns one combined `status`, rather
than requiring the caller to eyeball three separate tibbles.

### 2.6 Predictive Checks

Beyond calibration against the fitted data, prior and posterior predictive
checks simulate data *from* the model to check the priors are reasonable
before fitting, and that the fitted model reproduces the observed data's
statistical structure after fitting.

```{r probit-predictive, eval = FALSE}
prior_pred <- simulate_probit_prior_predictive(fit, n_states = 100L, seed = 123L)
compute_prior_predictive_status(prior_pred)
#> $summary$fraction_probability_lt_0.001
#> [1] 0.478
#> $summary$fraction_all_resistant
#> [1] 0.0857

ppc_stats <- compute_probit_ppc_statistics(
  fit, n_states = 100L, seed = 123L,
  statistics = c("marginal", "resistant_count", "pairwise")
)
compute_posterior_predictive_status(ppc_stats)
#> $status
#> [1] "pass"
#> $family_status$marginal$status
#> [1] "ok"
#> $family_status$pairwise$status
#> [1] "ok"
```

`plot_probit_diagnostics(fit, output_dir, experiment_id, pathogen)` (needs
`bayesplot`) writes trace/rank/pair plots for the monitored parameters to a
PDF, and `plot_probit_posterior_predictive_checks()` visualises the
`ppc_stats` tables -- both are for interactive review, not part of the
scripted pipeline.

### 2.7 Unified Dispatcher

`estimate_resistance_profiles()` wraps both pathways behind one interface --
switch pathway with `method`:

```{r probit-dispatcher, eval = FALSE}
# Pathway 1 (convex)
estimate_resistance_profiles(data = marginals, method = "convex", panel_map = panel_map)

# Pathway 2 (Bayesian) -- runs steps 2.1-2.4 above in one call
estimate_resistance_profiles(
  data = event_class_data, method = "bayesian",
  class_cols = class_cols, fixed_effects = c("Age_normalised", "gender"),
  random_effects = c("center_name"), pathogen = "Klebsiella pneumoniae",
  outcome_col = "final_outcome", residual_structure = "identity",
  sampler_config = list(chains = 2, iter_warmup = 200, iter_sampling = 200, seed = 123)
)
#> [estimate_resistance_profiles] Pathway 2 complete.
#> $profiles      # same shape as aggregate_profiles_for_daly() output above
#> $eligibility
#> $diagnostics
#> $fitted_models
#> $config_used
```

Both pathways' `profiles`/`agg` output are the deliverable of this vignette
-- downstream burden calculation (RR assignment, YLL/YLD) is not yet covered
here.

---

# Pipeline at a Glance

```
Preprocessing output (prep_* pipeline)
          │
          ├─────────────────────────────┬─────────────────────────────────────┐
          ▼ Option 1: Convex (Pathway 1) ▼                Option 2: Bayesian Probit (Pathway 2)
  compute_marginal_resistance()          fit_bayesian_multivariate_probit()
          │  (Step 1: marginals)                 │  (facility event-level AST + covariates)
          ▼                                       ▼
  compute_pairwise_coresistance()        compute_event_profile_probabilities()
          │  (Step 2: pairwise co-R)             │  (posterior draws -> profile probs)
          ▼                                       ▼
  compute_resistance_profiles()          aggregate_profiles_for_daly()
          │  (Step 3: QP -> profiles)            │  (+ validate_*_calibration(),
          │                                       │    predictive checks)
          │    Alternative entry (aggregate       │
          │    marginals): validate_aggregate_    │
          │    inputs() -> estimate_profiles_     │
          │    convex()                           │
          │                                       │
          │         (or, either pathway via one call:
          │          estimate_resistance_profiles(method = "convex" | "bayesian"))
          │                                       │
          └───────────────────┬───────────────────┘
                               ▼
                  Resistance profile probabilities (this vignette's scope ends here;
                  downstream burden calculation -- RR assignment, YLL/YLD -- not yet covered)
```

---

# Session Info

```{r session}
sessionInfo()
```
