---
title: "Getting Started with anumaan"
author: "anumaan Development Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with anumaan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
```

# Introduction

**anumaan** is an R package for preprocessing antimicrobial resistance (AMR) surveillance data. It standardizes messy hospital data into analysis-ready formats with proper normalization, classification, and quality control.

## Key Features

- Automated pipeline: process entire datasets with one function
- Organism normalization: handles typos, abbreviations, and variants
- Antibiotic classification: WHO AWaRe categories
- MDR/XDR classification: WHO/CDC-compliant definitions
- Event deduplication: creates unique infection episodes
- Data validation: comprehensive quality checks

## Installation

```{r, eval = FALSE}
# From source during development
devtools::install()
```

## Load the Package

```{r}
library(anumaan)
library(dplyr)
```

---

# Quick Start

## The Complete Pipeline

```{r, eval = FALSE}
library(readxl)
raw_data <- read_excel("your_data.xlsx")

result <- run_preprocess(
  data = raw_data,
  config = amr_config(
    hai_cutoff = 2,
    mdr_definition = "CDC",
    fuzzy_match = TRUE
  ),
  verbose = TRUE,
  validate = TRUE,
  generate_report = TRUE
)

clean_data <- result$data
summary(result)
```

## What the Pipeline Does

**Phase 1: Standardization**

- Maps column names automatically (fuzzy match)
- Normalizes organism and antibiotic names
- Standardizes specimen types, sex, and outcome values
- Parses dates and cleans R/S/I values

**Phase 2: Enrichment**

- Calculates age from DOB
- Computes length of stay
- Classifies HAI vs CAI
- Infers missing hospital departments

**Phase 3: Derivation**

- Creates event IDs (infection episode deduplication)
- Classifies MDR/XDR
- Flags contaminants
- Calculates polymicrobial weights
- Maps to GBD pathogens

---

# Sample Data Example

```{r}
sample_data <- data.frame(
  PatientID = c(1, 1, 2, 3),
  Age = c(65, 65, 34, 78),
  Gender = c("M", "M", "F", "M"),
  SampleDate = as.Date(c("2024-01-15", "2024-01-15", "2024-01-20", "2024-02-01")),
  Specimen = c("Blood", "Blood", "Urine", "Blood"),
  Organism = c("E. coli", "S. aureus", "E. coli", "K. pneumoniae"),
  Antibiotic = c("Ampicillin", "Oxacillin", "Ciprofloxacin", "Meropenem"),
  Result = c("R", "S", "S", "R")
)

result <- run_preprocess(sample_data, verbose = FALSE)

head(result$data)
```

---

# Using Individual Functions

## Organism Normalization

```{r}
data <- data.frame(
  organism_name = c("E. coli", "ACINETOBACTER BAUMANNII", "K. pneumoniae")
)

normalized <- prep_standardize_organisms(data, organism_col = "organism_name")
normalized[, c("organism_name", "organism_normalized")]
```

## Antibiotic Normalization

```{r}
data <- data.frame(
  antibiotic_name = c("Ampicillin", "Meropenem", "Ciprofloxacin")
)

normalized <- prep_standardize_antibiotics(data, antibiotic_col = "antibiotic_name")
normalized[, c("antibiotic_name", "antibiotic_normalized", "antibiotic_class")]
```

## MDR Classification

```{r, eval = FALSE}
# Requires normalized organism and antibiotic data with antibiotic_value column
classified <- prep_classify_mdr(
  processed_data,
  mdr_definition = "CDC"
)
table(classified$mdr_status)
```

## Wide-to-Long Conversion

```{r, eval = FALSE}
# Convert wide format (one column per antibiotic) to long format
long_data <- prep_pivot_ast_wide_to_long(
  wide_data,
  id_cols = c("PatientID", "Age", "Gender")
)
```

---

# Configuration Options

```{r}
config <- amr_config(
  fuzzy_match = TRUE,
  hai_cutoff = 2,
  event_gap_days = 14,
  mortality_window = 14,
  age_bins = "GBD_standard",
  mdr_definition = "CDC",
  xdr_definition = "CDC",
  intermediate_as_resistant = TRUE,
  strict_validation = FALSE
)
```

---

# Data Validation

```{r, eval = FALSE}
validation <- validate_required_fields(
  raw_data,
  required_cols = c("PatientID", "Organism", "Antibiotic", "Result"),
  min_completeness = 0.8
)
validation$valid
validation$messages

quality <- validate_data_quality(result$data)
```

---

# Partial Pipeline

```{r, eval = FALSE}
# Only standardization
result <- run_preprocess(raw_data, phases = "standardize", verbose = TRUE)

# Standardization + Enrichment
result <- run_preprocess(raw_data, phases = c("standardize", "enrich"), verbose = TRUE)
```

---

# Working with Results

```{r}
# Main processed data
clean_data <- result$data

# Configuration used
config_used <- result$config

# Processing log
processing_log <- result$log
```

## Export Results

```{r, eval = FALSE}
write.csv(result$data, "processed_amr_data.csv", row.names = FALSE)
saveRDS(result, "preprocessing_result.rds")
export_report(result$report, "report.html", format = "html")
```

---

# Common Analysis Tasks

## Resistance Rates

```{r}
result$data %>%
  filter(!is.na(antibiotic_value)) %>%
  summarise(
    total = n(),
    resistant = sum(antibiotic_value == "R"),
    resistance_rate = round(100 * resistant / total, 1)
  )
```

## MDR Prevalence

```{r}
if ("mdr_status" %in% names(result$data)) {
  result$data %>%
    group_by(organism_normalized) %>%
    summarise(
      total = n(),
      mdr = sum(mdr_status == "MDR", na.rm = TRUE),
      mdr_rate = round(100 * mdr / total, 1)
    )
}
```

---

# Troubleshooting

**Column not found**: The pipeline auto-detects common variants via fuzzy matching. Check column names with `names(raw_data)` and pass explicit mappings via `amr_config(column_mappings = list(...))` if needed.

**Organism not normalizing**: Organisms absent from the reference database pass through unchanged. Check coverage with `table(result$data$organism_normalized)`.

**Too many/few events**: Adjust the gap window with `amr_config(event_gap_days = 7)` (more events) or `amr_config(event_gap_days = 30)` (fewer events).

---

# Session Info

```{r}
sessionInfo()
```
